A Process for Emulating Comparative Oncology Trials with Real-world Evidence Studies (ENCORE): Process Development and Methodological Considerations for Oncology Real-World Data

### 2025-07-06 manuscript/manuscript.qmd
moving Sebastian to second author position
COTA review round 2

**Authors**: Janick Weberpals^1^, Kenneth L. Kehl^2^, Donna R. Rivera^3^, Pallavi Mishra-Kalyani^3^, Catherine C. Lerro^3^, Erin Larkins^4^, Sonia Singh^4^, Preeti Narayan^4^, Richard Curley^4^, Paul Kluetz^4^, Georg Hahn^1^, Priyanka Anand^1^, Yanina Natanzon^5^, Andrew J. Belli^6^, Ching-Kun Wang^6^, Jenna Collins^7^**,** Jonathan Kish^7^**,** Janet Espirito^8^, Nicholas J Robert^8^, Robert J. Glynn^1^, Sebastian Schneeweiss^1^, Shirley V. Wang^1^

@@

To formally compare treatment effects between RCTs and their respective emulations, we will adapt the approach of the RCT-DUPLICATE project.[@franklin2020nonrandomized; @wang2023emulation] That is, for the primary endpoint of interest (HR for OS and corresponding 95% confidence intervals), we will derive three qualitative agreement metrics: statistical agreement, estimate agreement and agreement based on the standardized mean difference (SMD). Examples are illustrated @tbl-metrics.

@@

The RCT DUPLICATE initiative[@wang2023emulation; @franklin2020nonrandomized] identified multiple emulation challenges that may be similar in ENCORE. One particular aspect that we may not always be able to emulate is the exact distribution of patient characteristics of the trial population, or that the comparability between these two populations may not be known. This can be due to the lack of RWD granularity and comprehensiveness to emulate relevant eligibility criteria or the fact that large pivotal trials in oncology are typically conducted in multiple countries worldwide (all considered databases reflect the US only). This may be a factor especially given that the pathophysiology, prognosis, and factors that drive heterogeneous treatment outcomes of certain cancers differ between countries. For example, some cancer indications (e.g., GI cancers) or some genetic mutations (e.g., EGFR mutations in lung adenocarcinoma) are much more prevalent in certain geographic regions compared to the US.

>>>>

**Authors**: Janick Weberpals^1^, Sebastian Schneeweiss^1^, Kenneth L. Kehl^2^, Donna R. Rivera^3^, Pallavi Mishra-Kalyani^3^, Catherine C. Lerro^3^, Erin Larkins^4^, Sonia Singh^4^, Preeti Narayan^4^, Richard Curley^4^, Paul Kluetz^4^, Georg Hahn^1^, Priyanka Anand^1^, Yanina Natanzon^5^, Andrew J. Belli^6^, Ching-Kun Wang^6^, Jenna Collins^7^**,** Jonathan Kish^7^**,** Janet Espirito^8^, Nicholas J Robert^8^, Robert J. Glynn^1^, Shirley V. Wang^1^

@@

To formally compare treatment effects between RCTs and their respective emulations, we will adapt the approach of the RCT-DUPLICATE project.[@franklin2020nonrandomized; @wang2023emulation] That is, for the primary endpoint of interest (HR for OS and corresponding 95% confidence intervals), we will derive three qualitative agreement metrics: statistical agreement, estimate agreement and agreement based on the standardized mean difference (SMD). Examples are illustrated in @tbl-metrics.

@@

The RCT DUPLICATE initiative[@wang2023emulation; @franklin2020nonrandomized] identified multiple emulation challenges that may be similar in ENCORE. One particular aspect that we may not always be able to emulate is the exact distribution of patient characteristics of the trial population, or that the comparability between these two populations may not be known. This can be due to the lack of RWD granularity and comprehensiveness to emulate relevant eligibility criteria or the fact that large pivotal trials in oncology are typically conducted in multiple countries worldwide (all considered databases reflect the US only). This may be a factor especially given that the pathophysiology, prognosis, and factors that drive heterogeneous treatment outcomes of certain cancers differ between countries. For example, some cancer indications (e.g., GI cancers) or some genetic mutations (e.g., EGFR mutations in lung adenocarcinoma) are much more prevalent in certain geographic regions and populations compared to the US.
